---
title: "Proposal"
output: html_document
---

## Section 2 - Data
```{r load-data, message=FALSE}
library(tidyverse)
library(openintro)

# Load the data
write.csv(loans_full_schema, "data/loans_data.csv", row.names = FALSE)

```
```{r glimpse-data}
glimpse(loans_data)
```

The dataset contains `r nrow(loans_data)` observations and `r ncol(loans_data)` variables.

## Section 3 - Data Analysis Plan

### Research Question
[What factors contribute the most to a borrowers interest rate?]

### Variables

**Outcome variable (Y):**
- `interest_rate` - The interest rate assigned to the loan (continuous, numeric)

**Predictor variables (X):**
- `loan_amount` - Amount borrowed (continuous)
- `annual_income` - Borrower's income (continuous)
- `grade` - Loan grade (categorical: A, B, C, D, E, F, G)
- `term` - Loan term length (categorical: 36 or 60 months)

### Comparison Groups
[If applicable - for example: "We will compare loans with different grades (A vs B vs C, etc.) to see how grade affects interest rate"]

### Preliminary Exploratory Data Analysis
```{r summary-stats}
# Summary statistics for key variables
loans_data %>%
  select(interest_rate, loan_amount, annual_income) %>%
  summary()
```
```{r visualizations}
# Distribution of interest rates
ggplot(loans_data, aes(x = interest_rate)) +
  geom_histogram(bins = 30, fill = "steelblue") +
  labs(title = "Distribution of Interest Rates",
       x = "Interest Rate (%)",
       y = "Count")

# Relationship between loan grade and interest rate
ggplot(loans_data, aes(x = grade, y = interest_rate)) +
  geom_boxplot(fill = "steelblue") +
  labs(title = "Interest Rate by Loan Grade",
       x = "Loan Grade",
       y = "Interest Rate (%)")
```

**Insights from EDA:**
- [Describe what you observe - e.g., "Interest rates range from X% to Y%, with a median of Z%"]
- [Note any patterns - e.g., "Higher loan grades (A, B) tend to have lower interest rates"]

### Proposed Methods

**Primary method:** Multiple linear regression
- Model interest rate as a function of loan amount, income, grade, and term
- This will allow us to see which factors most strongly predict interest rates

**Secondary methods:**
- ANOVA to compare interest rates across loan grades
- Correlation analysis between numeric variables

### Expected Results to Support Hypothesis

To support our hypothesis [state your hypothesis], we would need:
- Statistically significant coefficients (p < 0.05) for our key predictors
-  A negative coefficient for annual_income, suggesting higher income leads to lower rates
-  Significant differences between loan grades in ANOVA test
- Model with reasonable R² value indicating good fit